---
title: "P8105 -- Homework 1"
output: github_document
---

```{r library, message = FALSE}
library(tidyverse)
```

# Problem 1

Load the `penguins` dataset.

```{r load_data}
data("penguins", package = "palmerpenguins")
```

## Brief summary

Look at the dataset using `view()`.

```{r, eval = FALSE}
View(penguins)
```

Look at a brief summary of the dataset using `summary()`.

```{r summary}
summary(penguins)
```

Discussion:  

*  The variables involved in the `penguins` dataset are **species**, **island**, **bill_length_mm**, **bill_depth_mm**, **flipper_length_mm**, **body_mass_g**, **sex**, and **year**. For a brief summary of this dataset:
    * There are 3 speicies of penguins, including Adelie, Chinstrap, and Gentoo.
    * The penguins come from 3 unique islands, including Biscoe, Dream, and Torgersen.
    * The penguins in this dataset have bill lengths ranging from `r min(penguins$bill_length_mm, na.rm = TRUE) `mm to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm, bill depths ranging from `r min(penguins$bill_depth_mm, na.rm = TRUE)` mm to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm, flipper lengths ranging from `r min(penguins$flipper_length_mm, na.rm = TRUE)` mm to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm, and body masses ranging from `r min(penguins$body_mass_g, na.rm = TRUE)` g to `r max(penguins$body_mass_g, na.rm = TRUE)` g.
    * There are 165 female and 168 male penguins in the dataset.
    * All of the penguins were born between 2007 and 2009.  
*  The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.  
*  The mean flipper length for the penguins in the dataset is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

## Scatterplot

Now I will make a scatterplot of `flipper_length_mm` on y-axis and `bill_length_mm` on x-axis, with colored points by `species`.

```{r scatterplot, warning = FALSE}
ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
  ) +
  geom_point()
```

Export the scatterplot to the working directory.

```{r export, warning = FALSE}
ggsave("scatterplot_flipper_bill_length_by_species.pdf", height = 4, width = 6)
```

# Problem 2

## Create dataframe

```{r df}
p2_df =
  tibble(
    var_norm = rnorm(n = 10),
    var_logical = var_norm > 0,
    var_character = c("alopecia", "anxiety", "asthma", "chills", "deafness", "fever", "headache", "infection", "nausea", "rash"),
    var_factor = factor(c("severe", "severe", "mild", "moderate", "mild", "severe", "moderate", "mild", "mild", "mild"))
  )
```

## Take the mean

```{r mean, warning = FALSE}
mean(pull(p2_df, var_norm))
mean(pull(p2_df, var_logical))
mean(pull(p2_df, var_character))
mean(pull(p2_df, var_factor))
```

Taking the mean of numeric (`var_norm`) and logical (`var_logical`) variables works, but when taking the mean of character (`var_character`) and factor (`var_factor`) variables, it does not work.

## Convert variable type: to numeric

I will achieve it by `as.numeric` function.

```{r conversion, eval = FALSE}
as.numeric(pull(p2_df, var_logical))
as.numeric(pull(p2_df, var_character))
as.numeric(pull(p2_df, var_factor))
```

I could successfully convert the logical variable into numeric. However, I failed to convert the character into numeric. This happens because R can treat the logical "TRUE" and "FALSE", as "1" and "0", respectively, by default, but the character and factor cannot be treated as number (though factor can be re-coded by numbers to represent the "levels").  
As logical variable can be converted into numeric, character and factor cannot, it helps to explain what happens when trying to take the mean.

The factor can be successfully converted to numeric, taking values of 1 for `"mild"`, 2 for `"moderate"`, and 3 for `"severe"`. Numeric values are assigned to the factor levels, which are alphabetical unless otherwise specified (which is also reasonable in this context).
